DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car

### 2018-04-20 paper/intro.tex
intro revision

% broad context:
% - advance in ai sparked interests in the robotics application, such as
%   self-driving cars.
% - in particular, deep neural network models are increasingly used
%   for perception and control of a vehicle. say. AI workloads.
%
%
Autonomous cars have been a topic of increasing interest in recent
years as many companies are actively developing related hardware
and software technologies toward fully autonomous driving capability with
no human intervention. Deep neural networks (DNNs) have been
successfully applied in various perception and control tasks in
recent years.  They are important workloads for autonomous vehicles
as well. For example, Tesla Model S was known to use a specialized
chip (MobileEye EyeQ), which used a deep neural network for vision-based
real-time obstacle detection and avoidance. More recently, researchers
are investigating DNN based end-to-end control of
cars~\cite{Bojarski2016} and other robots. It is expected that more
DNN based artificial intelligence (AI) workloads may be used in future
autonomous vehicles.

% big problem
Executing these AI workloads on an embedded computing platform 
poses several additional challenges. First, many AI workloads in vehicles 
are computationally demanding and have strict real-time requirements. 
For example, latency in a vision-based object
detection task may be directly linked to the safety of the vehicle. This
requires a high computing capacity as well as the means to guaranteeing
the timings. On the other hand, the computing hardware platform must
also satisfy cost, size, weight, and power constraints, which require a
highly efficient computing platform. These two conflicting
requirements complicate the platform selection process as observed in
~\cite{Otterness2017}.

@@

% DeepPicar introduction
In this paper, we present DeepPicar, a low-cost autonomous car
testbed for research. From a hardware perspective,
DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
computer, a web camera and a RC car, all of which are affordable
components (less than \$100 in total).
The DeepPicar, however, employs a state-of-the-art AI
technology, namely end-to-end deep learning based real-time control,
which utilizes a deep convolutional neural network (CNN).
The CNN receives an image frame from a single forward
looking camera as input and generates a predicted steering angle
value as output at each control period in \emph{real-time}.
The CNN has 9 layers, about 27 million connections
and 250 thousand parameters (weights).
DeepPicar's CNN architecture is identical to that of NVIDIA's
real-sized self-driving car, called DAVE-2~\cite{Bojarski2016}, which
drove on public roads without human driver's intervention while only
using the CNN.
%% Note that, DAVE-2 run on a significantly more powerful computing
%% platform (Drive PX~\cite{drivepx}).

@@

% what are our findings?
Using DeepPicar, we systematically analyze its real-time
capabilities in the context of end-to-end deep-learning based real-time
control, especially on real-time \emph{inferencing} of the CNN.
We also evaluated other, more powerful, embedded computing
platforms to better understand achievable real-time performance of
DeepPicar's deep-learning based control system and the impact of
computing hardware architectures.
From the systematic study, we want to answer the following questions:
(1) How well does the CNN based real-time inferencing task perform on
contemporary embedded multicore computing platforms? (2) How
susceptible is the CNN inferencing task to contention in shared
hardware resources (e.g., cache and DRAM) when multiple tasks/models
are consolidated? (3) Are the existing state-of-the-art shared resource
isolation techniques effective in protecting real-time performance of
the CNN inferencing task?

% Findings:
%
% - we find DNN processing is an ideal real-time workload from the
%   perspective of timing predictability as the amount of computation
%   needed is fixed at design time and doesn't change over input.
%
% - real-time processing of non-trivial DNN inferencing is feasible
%   on today's embedded multicore platform, even as inexpensive as
%   raspberry pi3.
%
% - all tested platforms have enough computing capacity to consolidate 
%   multiple DeepPicar's CNN workloads or other compute intensive tasks
%   simultaneously.
%
% - however, when consolidating multiple workloads shared resource
%   contention is an important issue. we carefully analyzed
%   sensitivity of DeepPicar's workload with respect to shared
%   hardware resources.
%
% - in particular, we find DNN processing is highly sensitive to
%   memory performance. performance severely affected when memory
%   bandwidth is contented.
%
% - however, we find that dnn processing is not sensitive to cache
%   space. we find applying cache partitioning does not help improve
%   isolation of DNN workload when memory bandwidth is contended.
Our observations are as follows.
First, we find that real-time processing of the CNN inferencing is
feasible on contemporary embedded computing platforms, even one   
as inexpensive as the Raspberry Pi 3. Second, while consolidating
multiple CNN models and task are feasible but the impact of shared
resource contention can be catastrophic in the worst-case as we
observe up to 11.6X slowdown even when the CNN model was given a
dedicated core. Third, using the state-of-the-art OS-level cache
partitioning and memory bandwidth control
mechanisms~\cite{yun2014rtas,Yun2013}, we find that cache space has a
minimal performance impact to the CNN workload while memory
performance has a significant impact.

>>>>

% broad context:
% - advance in ai sparked interests in the robotics application, such as
%   self-driving cars.
% - in particular, deep neural network models are increasingly used
%   for perception and control of a vehicle. say. AI workloads.
%
%
Autonomous cars have been a topic of increasing interest in recent
years as many companies are actively developing related hardware
and software technologies toward fully autonomous driving capability with
no human intervention. Deep neural networks (DNNs) have been
successfully applied in various perception and control tasks in
recent years.  They are important workloads for autonomous vehicles
as well. For example, Tesla Model S was known to use a specialized
chip (MobileEye EyeQ), which used a vision-based real-time obstacle
detection system based on a DNN. More recently, researchers 
are investigating DNN based end-to-end real-time control for
robotics applications~\cite{Bojarski2016,Levine2016}. It is expected
that more DNN based artificial intelligence (AI) workloads may be
used in future autonomous vehicles.

% big problem
Executing these AI workloads on an embedded computing platform 
poses several additional challenges. First, many AI workloads,
especially those in vehicles, are computationally demanding and have
strict real-time requirements.
For example, computing latency in a vision-based object
detection task may be directly linked to the safety of the vehicle. This
requires a high computing capacity as well as the means to guaranteeing
the timings. On the other hand, the computing hardware platform must
also satisfy cost, size, weight, and power constraints, which require a
highly efficient computing platform. These two conflicting
requirements complicate the platform selection process as observed in
~\cite{Otterness2017}.

@@

% DeepPicar introduction
In this paper, we present DeepPicar, a low-cost autonomous car
testbed for research. From a hardware perspective,
DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
computer, a web camera and a small RC car, all of which are affordable
components (less than \$100 in total).
The DeepPicar, however, employs a state-of-the-art AI
technology, namely end-to-end deep learning based real-time control,
which utilizes a deep convolutional neural network (CNN).
The CNN receives an image frame from a single forward
looking camera as input and generates a predicted steering angle
value as output at each control period in \emph{real-time}.
The CNN has 9 layers, about 27 million connections
and 250 thousand parameters (weights).
DeepPicar's CNN architecture is identical to that of NVIDIA's
real-sized self-driving car, called DAVE-2~\cite{Bojarski2016}, which
drove on public roads without human driver's intervention while only
using the CNN.
%% Note that, DAVE-2 run on a significantly more powerful computing
%% platform (Drive PX~\cite{drivepx}).

@@

% what are our findings?
Using DeepPicar, we systematically analyze its real-time
capabilities in the context of end-to-end deep-learning based real-time
control, especially on real-time \emph{inferencing} of the CNN.
We also evaluate other, more powerful, embedded computing
platforms to better understand achievable real-time performance of
DeepPicar's CNN based control system and the performance impact of
computing hardware architectures.
From the systematic study, we want to answer the following questions:
(1) How well does the CNN based real-time inferencing task perform on
contemporary embedded multicore computing platforms? (2) How
susceptible is the CNN inferencing task to contention in shared
hardware resources (e.g., cache and DRAM) when multiple tasks/models
are consolidated? (3) Are the existing state-of-the-art shared resource
isolation techniques effective in protecting real-time performance of
the CNN inferencing task?

% Findings:
%
% - we find DNN processing is an ideal real-time workload from the
%   perspective of timing predictability as the amount of computation
%   needed is fixed at design time and doesn't change over input.
%
% - real-time processing of non-trivial DNN inferencing is feasible
%   on today's embedded multicore platform, even as inexpensive as
%   raspberry pi3.
%
% - all tested platforms have enough computing capacity to consolidate 
%   multiple DeepPicar's CNN workloads or other compute intensive tasks
%   simultaneously.
%
% - however, when consolidating multiple workloads shared resource
%   contention is an important issue. we carefully analyzed
%   sensitivity of DeepPicar's workload with respect to shared
%   hardware resources.
%
% - in particular, we find DNN processing is highly sensitive to
%   memory performance. performance severely affected when memory
%   bandwidth is contented.
%
% - however, we find that dnn processing is not sensitive to cache
%   space. we find applying cache partitioning does not help improve
%   isolation of DNN workload when memory bandwidth is contended.
Our main observations are as follows.
First, we find that real-time processing of the CNN inferencing is
feasible on contemporary embedded computing platforms, even one   
as inexpensive as the Raspberry Pi 3. Second, while consolidating
additional CNN models and tasks on a single multicore platform is
feasible but the impact of shared resource contention can be
catastrophic---we observe up to 11.6X slowdown even when the CNN model
was given a dedicated core.
Third, our evaluation of existing cache
partitioning and memory bandwidth throttling
techniques~\cite{yun2014rtas,Yun2013} shows that cache partitioning is
not effective in protecting the CNN inferencing task while memory
bandwidth throttling is quite effective.
